# ESM 232 Assignment 3: Almond Yield & Profit

---
author: "Kelsey Warren and Taylor Cook"
date: "2025-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages

```{r}
library(tidyverse)
library(here)
library(purrr)
```

#1. Develop a profit model for your almond yield

- Use your imagination to think about what would make sense
- Make sure you include parameters
- You could assume a baseline profit and then adjust according to the anomaly


```{r}
#load almond yield data
clim <- read.table(here("data/clim.txt"), header = TRUE, sep = "", strip.white = TRUE)

source(here("~/Bren MESM/ESM 232 ENV Modeling/ESM232_EnvModeling/R/almondyield.R"))
source(here("~/Bren MESM/ESM 232 ENV Modeling/ESM232_EnvModeling/R/almond_profit.R"))

```


# 2. Apply the profit model to your almond yield model

- There are many ways to combine the almond yield and profit functions; you can have the profit function "call"/use the almond yield function; or create a wrapper function that calls them in sequence (first the almond yield and then the profit function)



```{r}
# IDK IF THIS PART IS RIGHT!! Need to dig into it more

# Wrapper function to calculate almond profit by first calculating the yield anomaly

almond_profit_wrapper <- function(df, month, acres = 500, baseline_profit = 4000, 
                                  price_per_ton = 6000, base_water_cost = 200) {
  
  # Step 1: Calculate the yield anomaly using the almondyield function
  yield_anomaly <- almondyield(year = unique(df$year), clim = df)
  
  # Step 2: Calculate the profit using the almond_profit function
  profit <- almond_profit(df = df, month = month, acres = acres, 
                          baseline_profit = baseline_profit, 
                          price_per_ton = price_per_ton, 
                          base_water_cost = base_water_cost)
  
  # Return the profit calculated
  return(profit)
}

```



# 3. Perform  a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

```{r}

```


# 4. Create 2 graphs
    - one that shows yield anomaly for each year, accounting for uncertainty in the parameters
    - one that show how yield anomaly in general varies with your parameters


```{r}

```




# 5. Write a short paragraph (in a Quatro document) to summarize you interpretation of your model results (e.g what do 'take away' from your model and sensitivity analysis)







