# ESM 232 Assignment 3: Almond Yield & Profit

---
author: "Kelsey Warren and Taylor Cook"
date: "2025-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages

```{r laod_packages}
library(tidyverse)
library(here)
library(purrr)
library(ggpubr)
```

# 1. Develop a profit model for your almond yield

-   Use your imagination to think about what would make sense
-   Make sure you include parameters
-   You could assume a baseline profit and then adjust according to the anomaly

```{r load_data}
#load almond yield data
clim <- read.table(here("data/clim.txt"), header = TRUE, sep = "", strip.white = TRUE)

# load functions to calculate almond yield, almond profit, and the wrapper function to bring them together
source(here("~/Bren MESM/ESM 232 ENV Modeling/ESM232_EnvModeling/R/almondyield.R"))
source(here("~/Bren MESM/ESM 232 ENV Modeling/ESM232_EnvModeling/R/almond_profit.R"))
source(here("~/Bren MESM/ESM 232 ENV Modeling/ESM232_EnvModeling/R/almond_wrapper.R"))

```

# 2. Apply the profit model to your almond yield model

-   There are many ways to combine the almond yield and profit functions; you can have the profit function "call"/use the almond yield function; or create a wrapper function that calls them in sequence (first the almond yield and then the profit function)

```{r wrapper}
# Wrapper function to calculate almond profit by first calculating the yield anomaly
# Note that our updated almondyield() function is designed to calcualte yield in a chosen single year at a time, rather than for the entire time period.
# This wrapper function was already loaded above, but showing the code here for the example year 2000
all_years = c(1989:2010)
almond_wrapper(years = all_years, clim = clim) 


# add results to a data frame in global environment

profit_df <- almond_wrapper(all_years, clim)

```


# 3. Perform a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

```{r sensitivity}
# NEED TO USE PMAP TO TEST PARAMETERS


# informal sensitivity analysis for all years
# generate sample data of uncertainty in parameter

# using 10% uncertainty in baseline_profit
nsamples = 200
deviation = 0.10
base_baseline_prof = 4000
baseline_profit <- runif(
  min = base_baseline_prof - deviation * base_baseline_prof,
  max = base_baseline_prof + deviation * base_baseline_prof, n = nsamples)
  
# using 10% uncertainty in price_per_ton
base_price = 6000
price_per_ton <- runif(
  min = base_price - deviation * base_price,
  max = base_price + deviation * base_price, n = nsamples)


# using 10% uncertainty in acres
base_acres <- 500
acres <- runif(nsamples, 
               min = base_acres - deviation * base_acres, 
               max = base_acres + deviation * base_acres)

# combine the uncertain parameter values into a data frame
params_u <- cbind.data.frame(baseline_profit, price_per_ton, acres)


# extract yield anomaly for all years
ya <- profit_df$yield_anomaly[profit_df$year == all_years]

# run the almond_profit() model using parameter uncertainty data generated above using pmap()
# takes function name and then names of all parameters that don't change
profit_sens_results <- params_u %>% 
  pmap(almond_profit,
    yield_anomaly = ya, 
    clim = clim,
    base_water_cost = 200)

#profit_sens_results[1]
#length(profit_sens_results)

# create a data frame to store the parameters and profit results
# but first unlist the profit results so they become a proper numeric vector
total_profit_df <- cbind.data.frame(
  total_profit = unlist(profit_sens_results),
  params_u
)

```

# 4. Create 2 graphs

```         
- one that shows yield anomaly for each year, accounting for uncertainty in the parameters
- one that show how yield anomaly in general varies with your parameters
```

```{r graphs}

# Add year information to `total_profit_df` by repeating the `year` values from `profit_df`
total_profit_df$year <- rep(profit_df$year, each = length(baseline_profit))


# Merge the two data frames by the year column
merged_df <- merge(profit_df, total_profit_df, by = "year")


merged_df$year <- as.numeric(as.character(merged_df$year))


# Plot 1: Yield Anomaly Over Time with uncertainty in profit parameter: price per ton
plot_1 <- ggplot(merged_df, aes(x = year, y = yield_anomaly, col = price_per_ton)) +
  # geom_point(cex = 2) +  
  geom_boxplot() +
  labs(y = "Yield Anomaly (ton/acre)", x = "Year") +
  scale_color_gradient(low = "blue", high = "red", name = "Price Per Ton ($)") +
  theme_minimal()
plot_1


# Plot 2: Yield over time with uncertainty by acres
plot_2 <- ggplot(merged_df, aes(x = factor(year), y = total_profit)) +
  geom_boxplot(fill = "lightgreen", outlier.colour = "gray50", outlier.alpha = 0.4) + 
  labs(y = "Total Profit ($)", x = "Year", title = "Total Profit Over Time with Uncertainty in Acres") +
  theme_minimal()
plot_2
# view plots side by side
ggarrange(plot_1, plot_2)

```

# 5. Write a short paragraph (in a Quarto document) to summarize you interpretation of your model results (e.g what do 'take away' from your model and sensitivity analysis)

The almond yield and profit model indicates how uncertainties in key parameters we chose, baseline profit and price per ton, influence almond production and total profit over time. The first plot, showing yield anomaly across years with uncertainty, demonstrates that small fluctuations in price per ton can significantly affect the yield, especially in years with extreme climate conditions. The color gradient helps visualize the sensitivity of yield anomalies to price changes, where a higher price per ton is associated with greater yield anomalies, suggesting more significant profitability during favorable climate years.

The second plot, depicting total profit over time with uncertainty, highlights the role of baseline profit in determining total profit variability. As expected, fluctuations in baseline profit are closely tied to the total profit outcomes, with higher baseline profits leading to greater total profit. The results from the sensitivity analysis show that while yield anomaly has the largest effect on profitability in certain years, baseline profit and price per ton both contribute to the overall variability. These findings suggest that understanding the potential fluctuations in key parameters is crucial for predicting the economic viability of almond farming under changing climate conditions and market prices.








