# ESM 232 Assignment 3: Almond Yield & Profit

### author: "Kelsey Warren and Taylor Cook"
### date: "2025-04-21"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages

```{r laod_packages}
library(tidyverse)
library(here)
library(purrr)
library(ggpubr)
```

# Develop a profit model for your almond yield

-   Use your imagination to think about what would make sense
-   Make sure you include parameters
-   You could assume a baseline profit and then adjust according to the anomaly

```{r load_data}
#load almond yield data
clim <- read.table(here("data/clim.txt"), header = TRUE, sep = "", strip.white = TRUE)

# load functions to calculate almond yield, almond profit, and the wrapper function to bring them together
source(here("~/Bren MESM/ESM 232 ENV Modeling/ESM232_EnvModeling/R/almondyield.R"))
source(here("~/Bren MESM/ESM 232 ENV Modeling/ESM232_EnvModeling/R/almond_profit.R"))
source(here("~/Bren MESM/ESM 232 ENV Modeling/ESM232_EnvModeling/R/almond_wrapper.R"))

```


# Sensitivity Analysis 

```{r}
# Generate the randomized inputs
set.seed(100)
# Number of samples (23 years * 100 iterations)
n_samples <- 2300
year <- rep(1988:2010, 100)  # 23 years * 100 iterations


# Randomize for precipitation (mean = 200mm, sd = 40mm)
precip <- rnorm(n = n_samples, mean = 400, sd = 40)

# Randomize for minimum temperature (uniform distribution between 7 and 12°C)
min_temp <- runif(n = n_samples, min = 15, max = 30)

# Randomize for acres (size of the farm, uniform distribution between 300 and 600 acres)
acres <- runif(n = n_samples, min = 300, max = 600)

# Create a data frame with randomized parameters (year, precipitation, minimum temperature)
parms <- cbind.data.frame(year, precip, min_temp, acres)

# Function to calculate almond yield and profit using the randomized parameters
results <- parms %>% pmap(almond_wrapper)

# Transform the results into a data frame containing yield, profit, and year
output_df <- map_df(results, `[`, c('yield_anomaly', 'profit'))

# Combine the results with the original parameters (year, precip, min_temp)
final_df <- cbind.data.frame(output_df, parms)

# Check the structure of the data
head(final_df)

```



# Create 2 graphs


```{r graphs}


# Plot 1: Boxplot for Almond profit by year
plot_1 <- ggplot(final_df, aes(x = factor(year), y = profit)) +
  geom_boxplot(fill = "steelblue") +  
  labs(
    title = "Almond Profit by Year",
    x = "Year",
    y = "Profit ($)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))  # Rotate x-axis labels for readability
plot_1

# Plot 2: Boxplot for Almond Yield by Year
plot2 <- ggplot(final_df, aes(x = factor(year), y = yield_anomaly, group = year)) +
  geom_boxplot(fill = "seagreen", alpha = 0.6, outlier.shape = NA) +  # Adjust boxplot appearance
  scale_color_gradient(low = "purple", high = "orange", name = "Precipitation (mm)") +  # Precipitation gradient coloring
  labs(
    title = "Almond Yield Anomaly by Year",
    x = "Year",
    y = "Yield Anomaly (ton/acre)",
    caption = "Variation driven by precipitation & temperature uncertainty"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability

# Display the plot
plot2
```

# 5. Write a short paragraph (in a Quarto document) to summarize you interpretation of your model results (e.g what do 'take away' from your model and sensitivity analysis)

The almond yield and profit model demonstrates how uncertainties in key parameters, including precipitation, temperature, and acres, influence almond production and total profit over time. Plot 1, which shows the almond profit by year, illustrates the significant impact of variability in climate factors (precipitation and minimum temperature) as well as farm size (acres) on profit. The boxplot reveals that while profits remain relatively stable across years, the spread of profits widens in years with extreme climate conditions or larger farm sizes. This variability underscores the sensitivity of almond farming profitability to climate uncertainties, particularly in years with more extreme precipitation or temperature anomalies, as well as variations in farm size.

Plot 2, which depicts the almond yield anomaly by year, emphasizes how fluctuations in yield are influenced by climate variability and farm size. The boxplot reveals that yield anomalies exhibit considerable variation, with the spread of yields being wider during years with extreme climate conditions. The results indicate that variations in precipitation, temperature, and acres have a direct impact on almond yields, with years of unfavorable climate conditions leading to lower yields, while favorable conditions and larger farm sizes can lead to higher yields.









