---
---
---

# Assignment 8: Calibration Part 2

## Taylor Cook

### 2025-06-10

### Use your performance metric that you developed in Calibration Part 1

```{r}
library(tidyverse)
library(here)
library(sensitivity)
library(lubridate)

msage = readRDS(here("data/msage.RDS"))

msagel = msage %>% gather(key="run",value="str", -date, -month, -day, -year, -wy,-obs)

source(here("R/combined_rmse.R"))
compute_all_metrics

```

#### Use the performance values as weights to come up with a max likelihood estimate (MLE) of

```{r}


# Apply your custom accuracy metric to each run (excluding date and obs columns)
res <- msage %>%
  select(-date, -month, -day, -year, -wy, -obs) %>%
  apply(2, compute_all_metrics, o = msage$obs, month = msage$month)


# Convert to data frame
results <- as.data.frame(matrix(unlist(res), byrow = TRUE, ncol = 1))  # Assuming your function returns 1 metric
colnames(results) <- "performance"
results$run <- colnames(msage %>% select(-date, -month, -day, -year, -wy, -obs))

#Nomralize performance as weights
# Normalize to [0, 1]
max_perf <- max(results$performance, na.rm = TRUE)
max_perf

min_perf <- min(results$performance, na.rm = TRUE)
min_perf

results <- results %>%
  mutate(wt = (performance - min_perf) / (max_perf - min_perf),
         wt = wt / sum(wt, na.rm = TRUE))  # Normalize so they sum to 1

#add weights to msagel
# Join weights with long-form streamflow data
msagel <- left_join(msagel, results, by = "run")

```

a)  daily streamflow

```{r}
# Compute the MLE for daily streamflow
mle_daily <- msagel %>%
  group_by(date) %>%
  summarise(mle_str = weighted.mean(str, w = wt, na.rm = TRUE)) %>%
  ungroup()
mle_daily

```

b)  another component of streamflow that is relevant given your performance metric (e.g August streamflow in the class example)

```{r}
# Compute the MLE for August streamflow
mle_august <- msagel %>%
  filter(month == 8) %>%
  group_by(date) %>%
  summarise(mle_str_august = weighted.mean(str, w = wt, na.rm = TRUE)) %>%
  ungroup()
mle_august
```

#### Graph MLE and observed for the streamflow component of interest

```{r}
# Extract observed streamflow
obs_plot <- msage %>%
  select(date, obs) %>%
  distinct()

# Plot MLE and observed streamflow
ggplot() +
  geom_line(data = obs_plot, aes(x = date, y = obs, color = "Observed"), size = 1) +
  geom_line(data = mle_daily, aes(x = date, y = mle_str, color = "MLE Daily"), size = 1) +
  labs(title = "Daily Streamflow: Observed vs MLE",
       y = "Streamflow (cfs)",
       x = "Date") +
  scale_color_manual(values = c("Observed" = "lightblue", "MLE Daily" = "darkred")) +
  theme_minimal()
```

#### Compute the correlations between observed and the MLE and observed flow for daily flow and your streamflow component of interest

```{r}
# Join observed with MLE for daily
obs_mle_daily <- left_join(obs_plot, mle_daily, by = "date")

# Join for August only
obs_mle_august <- left_join(obs_plot, mle_august, by = "date") %>%
  filter(month(date) == 8)

# Compute correlations
cor_daily <- cor(obs_mle_daily$obs, obs_mle_daily$mle_str, use = "complete.obs")
cor_daily
cor_august <- cor(obs_mle_august$obs, obs_mle_august$mle_str_august, use = "complete.obs")
cor_august

cat("Correlation between observed and MLE daily streamflow:", cor_daily, "\n")
cat("Correlation between observed and MLE August streamflow:", cor_august, "\n")
```
