# Assignment 5: Using Sobol with an ODE solver

## Taylor Cook

### 2025-05-12

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
library(sensitivity)
library(here)
library(ggpubr)
library(lhs)
library(purrr)
```

Part 1

Build this model (e.g add hunting to the lotvmodK.R),

**Some requirements/hints for your model**

*You should make sure that you don't hunt more prey than exist.*

*To ensure that you might also add a minimum prey population input that must be met before hunting is allowed.*

*Note you can make this as simple or as complex as you would like. You could represent hunting in a way that is similar to "harvesting" in the last assignment.*



```{r}
lotvmodK <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    dPrey <- r * Prey * (1 - Prey / K) - a * Prey * Pred
    dPred <- e * a * Prey * Pred - m * Pred
    
    # Hunting effect
    if (Prey > min_prey) {
      dPrey <- dPrey - h * Prey
    }
    
    list(c(dPrey, dPred))
  })
}




```









Part 2

Explore how different hunting levels and different minimum prey populations (before hunting is allowed) are likely to effect the stability of the populations of both predator and prey.
A key challenge is how you might want to define stability? It is up to you but you will need to write a sentence to explain why you chose the measure that you did. 
It could be something as simple as maintaining a population above some value 50 years into the future.  

Use this exploration to recommend a hunting target that will be sustainable (e.g leave you with a stable prey and predator population).

It is up to you how you "explore" hunting  - you can simply try different values of the parameters in your hunting model or do it more formally by running your model across a range of values. You could think about parameter interactions

You can assume the following are best guesses of key parameters

rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000,




